---
title: "Assignment2"
subtitle: "BCB 504 - Foundations of Data Visualization"
author: "Ronald Bentil"
date: "2023-01-30"
format: html
code-fold: true
code-tools: true
---

::: callout-note
I added additional content to your YAML header.  Your name, date, the course information are important and should be included in all future assignements.  Also added the code options.
:::

## My dataset

This is a data set collected during my Masters degree research. The data contains information on livestock and the ticks collected off them.  The ticks were subsequently identified morphologically and the pathogens that these ticks harbored were identified using PCR on extracted nucleic acid.

I'm putting this dataset together in response to [Assignment 2](https://professorpolymorphic.github.io/RobisonWebSite/BCB520/posts/A2-YourData/) from the [BCB 504](https://professorpolymorphic.github.io/RobisonWebSite/BCB520/BCB520.html) class.

## Data collection

The data were collected from two parts of Ghana (Greater Accra in the South and Upper East in the North). Sampling sites were carefully chosen based on known climate and environmental conditions that influence arthropods and their ability to transmit pathogens.

::: callout-note
I cleaned up some of the writing in the preceding sections.  Below, I renamed your data file to just `Ronald.xlsx`.  Your previous data file had parentheses in the name, which can cause major problems in automated data analyses and visualizations. I would also advise that you start thinking about a coherent naming convention for your data files.  Something that contains information about the data itself, such as `RB_MSThesisData_Date.xlsx` .  Also got rid of the `View()` statement as it was redundant with `glimpse()`.
:::

```{r}
#| echo: false
#| output: false

library(tidyverse)
library(readxl)

Results_For_Ronald_ <- read_excel("Ronald.xlsx")

```
### Data

The following shows the basic formatting of the data file using the `glimpse()` function:

```{r}

tickdata<-Results_For_Ronald_
glimpse(tickdata)

```

Note that some of the Attributes have `NA` entries.  This is a consequence of the data file, and I will fill in those entries and move the data file toward a [TidyData](https://towardsdatascience.com/what-is-tidy-data-d58bb9ad2458) format as I progress through the course.

Below is the data dictionary for my data file:

```{r}

 tickdata_attributes <- read_excel("TickData_Attributes.xlsx")
 options(knitr.kable.NA = '')
 knitr::kable(tickdata_attributes,caption='Data Dictionary for the data file Ronald.xlsx')

```

::: callout-note
Suppressed the NA results in your dictionairy with the `options()` code, and added a caption to your table. 
:::